---
title: "Using `rama`"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

# 1 Preambule

## 1.1 Installing `rama`

```{r}
# installed.packages("devtools")
# devtools::install_github("r-and-gama/rama")
```

## 1.2 Loading `rama`

```{r}
library(rama)
```

## 1.3 Configuring `rama`

```{r}
# Not sure we still need this...
```

# 2. Desiging experiments

We can make an `experiment` object either by reading an experiment from a GAML
file with the `load_experiment()` funtion or by building it from scratch with
the `experiment()` constructor.

## 2.1 Loading an experiment from a GAML file

Let's use the following `gaml` file:

```{r}
gaml_file <- "/Users/choisy/Dropbox/aaa/r-and-gama/rama/inst/examples/sir.gaml"
```

Loading the experiment `sir` from this file:

```{r}
exp1 <- load_experiment("sir", gaml_file)
```

which gives:

```{r}
exp1
```

An `experiment` is a simple data frame:

```{r}
class(exp1)
```

This implies that we will be able to use all the data frame methods to
manipulate experiments. More on that later.

## 2.2 Building an experiment from scratch

Alternatively, an experiment can be built from scratch with the `experiment`
constructor:

```{r}
exp2 <- experiment(
  data.frame(S0 = 999, I0 = 1, R0 = 0, beta = 1.5, gamma = .15),
  data.frame(S = 1, I = 1, R = 1),
  tmax = 1000,
  seed = 1,
  model = gaml_file
)
```

Which gives the same thing:

```{r}
exp2
```

Note that if you enter values of inequal lengths, it will automatically
complete:

```{r}
exp3 <- experiment(
  data.frame(S0 = c(900, 950, 999), # this is a data frame of 3 lines
              I0 = c(100, 50, 1),
              R0 = 0,
              beta = 1.5,
              gamma = .15),
  data.frame(S = 1, I = 1, R = 1),  # this is a data frame of 1 line
  tmax = 1000,                    
  seed = 1,                         
  model = gaml_file
)
```

Which gives:

```{r}
exp3
```

Note finally, that we can make smart use of the `expand.grid()` function of the
`base` package to efficiently generate a complete experimental design:

```{r}
exp4 <- experiment(
  expand.grid(S0 = c(900, 950, 999),
               I0 = c(100, 50, 1),
               R0 = 0,
               beta = 1.5,
               gamma = .15),
  data.frame(S = 1, I = 1, R = 1),
  tmax = 1000,
  seed = 1,
  model = gaml_file
)
```

Which gives:

```{r}
exp4
```

We'll see more about that in the next sections.

## 2.3 Building an experiment from a data frame

```{r}
df <- data.frame(
  S0 = c(700, 800, 900, 950, 999),
  R0 = 0,
  beta = 1.5,
  gamma = .15,
  S = 1,
  I = 1,
  tmax = 2500,
  seed = sample(5)
)
```

Constant population size of 1000:

```{r}
df$I0 <- 1000 - df$S0
```

```{r}
df
```

Let's now transform this data frame into an `experiment`:

```{r eval = F, include = F}
experiment(df)
```


# 3. Manipulating experiments

As mentioned above, since `experiment` objects are data frames, all the data
frame methods can be used to manipulate them, such as `nrow()` to know the
number of simulations in an experiment:

```{r}
nrow(exp4)
```


## 3.1 Replicating experiments

From the following experiment

```{r}
exp1
```

We can replicate:

```{r}
exp5 <- repl(exp1, 10)
```

Which gives:

```{r}
exp5
```

On which we could use different seeds for each simulation:

```{r}
exp5$seed <- sample(10)
```

Which gives:

```{r}
exp5
```

## 3.2 Seeing and changing the model path

An experiment is linked to a model on a GAMA file. To get the path of this file:

```{r}
model(exp1)
```

To change the path of this file:

```{r}
model(exp1) <- "new_path"
```

Let's check:

```{r}
model(exp1)
```


## 3.3 Extracting parameters values

```{r}
parameters(exp5)
```


## 3.4 Extracting observation rates

```{r}
observation(exp5)
```


## 3.5 `tmax` and `seed`

As seen above, `seed` can be extracted and changed with regular data frame
methods:

```{r}
exp5$seed
exp5$seed <- sample(10)
exp5
```

Same for `tmax`:

```{r}
exp5$tmax
exp5$tmax <- 3543
exp5
```

If you want to stop observing the variable R and observe the variable I every
2 steps, you'd do:

```{r}
exp5$r_R <- NULL
exp5$r_I <- 2
```

Which gives:

```{r}
exp5
```


# 4. Running experiments
